---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warnings = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# tidylfs

<!-- badges: start -->
[![R-CMD-check](https://github.com/wklimowicz/tidylfs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/wklimowicz/tidylfs/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Installation

You can install the development version of tidylfs from [GitHub](https://github.com/) with:

```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("wklimowicz/tidylfs")
```

## Quickstart

Copy the LFS `.sav` files into a directory (in my example `lfs_data_folder`). They should look like this:

```
.
└── lfs_data_folder/
    ├── 1992 Q1.sav
    ├── 1992 Q2.sav
    ├── ...
    └── 2020 Q2.sav
```

You can then run:

```{r eval=FALSE}
library(tidylfs)

# Converts `.sav` files to `.Rds` files, to save
# space and for quicker loading
lfs_convert(
  directory = "lfs_data_folder/",
  output_directory = "lfs_rds_folder/"
)

# Compiles into one file.
lfs <- lfs_compile(directory = "lfs_rds_folder/")
```

To view the variable mapping use `variable_mapping()`:

```{r}
variable_mapping(lfs)
```

Code to reproduce some official ONS publications is in the `vignette("stats-replication")`. For more information on the ONS variables, see the [LFS Guidance](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes/methodologies/labourforcesurveyuserguidance).

A core set of variables is included by default. See `R/mappings.R` for the complete set. To add more variables see `vignette("adding-variables")`.

```{r echo = F}
library(magrittr)

x <- tibble::tribble(
  ~`Variable Name`, ~`Definition`,
  "YEAR", "Year",
  "QUARTER", "Quarter",
  "SEX", "Sex",
  "GOVTOR", "Government Office Region",
  "AGE", "Age",
  "ETHNICITY", "Ethnicity",
  "FTPTWK", "Part-Time/Full-time Status",
  "EDAGE", "Age when completed full time education",
  "TTACHR", "Actual hours worked",
  "HIQUALD", "Highest Qualification",
  "DEGREE_SUBJECT", "Degree Subject",
  "OCCUPATION", "Occupation in main job",
  "PARENTAL_OCCUPATION", "Parental Occupation at 14"
)

x %>%
  kableExtra::kbl(format = "pipe")
```

