---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warnings = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# tidylfs

<!-- badges: start -->
[![R-CMD-check](https://github.com/wklimowicz/tidylfs/workflows/R-CMD-check/badge.svg)](https://github.com/wklimowicz/tidylfs/actions)
<!-- badges: end -->

## Installation

You can install the development version of tidylfs from [GitHub](https://github.com/) with:

```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("wklimowicz/tidylfs")
```

## Quickstart

Copy the LFS `.sav` files into a directory (in my example `lfs_data_folder`). They should look like this:

```
.
└── lfs_data_folder/
    ├── 1992 Q1.sav
    ├── 1992 Q2.sav
    ├── ...
    └── 2020 Q2.sav
```

You can then run:

```{r eval=FALSE}
library(tidylfs)

# Converts `.sav` files to `.Rds` files, to save
# space and for quicker loading
lfs_convert(lfs_directory = "lfs_data_folder/",
            output_directory = "lfs_rds_folder/")

# Compiles into one file.
lfs <- lfs_compile(lfs_directory = "lfs_rds_folder/")
```

To reproduce official ONS publications, such as

* Unemployent [UNEM01 NSA](https://www.ons.gov.uk/employmentandlabourmarket/peoplenotinwork/unemployment/datasets/unemploymentbyageanddurationnotseasonallyadjustedunem01nsa)
```{r eval=FALSE}
lfs %>%
    lfs_summarise_unemployment(QUARTER)
```

* Salary by occupation [EARN06](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/earningsandworkinghours/datasets/grossweeklyearningsbyoccupationearn06)
```{r eval=FALSE}
lfs %>%
    dplyr::filter(
      !is.na(OCCUPATION_MAJOR), # Filter out NA's
      FTPTWK == "Full-time" # Take only full-time employees
    ) %>%
    lfs_summarise_salary(QUARTER, OCCUPATION_MAJOR)
```

Extending them is easy:

* Unemployment by quarter, sex, and age category
```{r eval=FALSE}
lfs %>%
  lfs_summarise_unemployment(QUARTER, SEX, AGES)
```

For more information on the ONS variables, see the [LFS Guidance](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes/methodologies/labourforcesurveyuserguidance).

Some of the variables included by default are:
```{r echo = F}

library(magrittr)
x <- tibble::tribble( ~`Variable Name`, ~`Definition`,
"YEAR", "Year",
"QUARTER", "Quarter",
"SEX", "Sex",
"GOVTOR", "Government Office Region",
"AGE", "Age",
"FTPTWK", "Part-Time/Full-time Status",
"EDAGE", "Age when completed full time education",
"TTACHR", "Actual hours worked",
"UNION", "In union?",
"HIQUALD", "Highest Qualification",
"DEGREE_SUBJECT", "Degree Subject",
"OCCUPATION", "Occupation in main job",
"PARENTAL_OCCUPATION", "Parental Occupation at 14",
"ETHNICITY", "Ethnicity"
)

x %>%
  kableExtra::kbl(format = "pipe")

```

# Using the data across multiple projects

To avoid storing multiple copies of the compiled dataset, you can set the `DATA_DIRECTORY` environment variable before compiling. This can be done on the system or in `.Rprofile`:

```{r eval = FALSE}
Sys.setenv(DATA_DIRECTORY = "path/to/folder")
```
When compiling the dataset include the save_to_folder option as `TRUE`:

```{r eval = FALSE}
lfs_compile(lfs_directory = "lfs_rds_folder/", save_to_folder = TRUE)
```
The compiled dataset will be saved to the directory as an fst file, and you can load it from anywhere using this command:

```{r eval=FALSE}
lfs <- lfs_load(data.table = FALSE)
```
