---
title: "Adding your own Variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adding your own Variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

# Overview

* `tidylfs` is designed to be easily extended. Adding variables and summaries is a few steps:
    * Find the relevant variables in the [ONS LFS User guidance](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes/methodologies/labourforcesurveyuserguidance).
    * Add them to the function below.
    * Load the function, and pass it as an argument to `lfs_compile`.

* For example, to add marital status:

```{r}
user_extra_mappings <- function(lfs_file_column_names) {

  custom_variables <- tibble::tribble(
    ~lfs_name,       ~new_name,     ~type,
    "MARSEX6", "SEX_AND_MARITAL_STATUS", "factor"
    )

  return(custom_variables)
}

lfs_compile("your_lfs_rds_directory", extra_mappings = user_extra_mappings)

```

* To add a variable that changes over time, use `pick_var`. As an example, degree class is `DEGCLS7` from 2007 onwards, and `DEGCLS` before then. This code imports those as a factor (retaining the SPSS labels), and calls the new column `DEGREE_CLASS`.

```{r}
user_extra_mappings <- function(lfs_file_column_names) {

  # This will pick DEGCLS7 if it exists in the quarterly file,
  # if not then DEGCLS, otherwise it'll return NA.
  degree_class <- pick_var(c("DEGCLS7", "DEGCLS"), lfs_file_column_names)

  custom_variables <- tibble::tribble(
    ~lfs_name,       ~new_name,     ~type,
    "MARSEX6", "SEX_AND_MARITAL_STATUS", "factor",
    degree_class, "DEGREE_CLASS", "factor"
    )

  return(custom_variables)
}

lfs_compile("your_lfs_rds_directory", extra_mappings = user_extra_mappings)

```
